---
title: "MEPS"
output: html_document
---

# MEPS Documentation

Notes:

- **Size**
  - Number of MEPS participants in 2013: 36,940 ([source](http://meps.ahrq.gov/mepsweb/data_stats/download_data/pufs/h163/h163doc.shtml) see "Data File Information")
- **Design**
  - Each year, a new "panel" of households is selected for the survey:
    - The households for each panel are selected from respondents to the National Health Interview Survey
    - The data collection for each year takes 2 and a half years, and is done in 5 rounds of interviews
  - See [here] (http://meps.ahrq.gov/survey_comp/hc_data_collection.jsp) for more details
  
  - 2004 Design summary ([source](https://apha.confex.com/apha/135am/recordingredirect.cgi/id/18758))
    - PSU’s grouped into 237 design strata
    – 358 PSU’s sampled across strata
    - Second Stage Units (SSU’s)
      – Clusters of housing units
      – Oversample of SSU’s with large Black/Hispanic populations
    - MEPS based on subsample of about 200 PSU’s from NHIS
    - Person weights account for:
      - oversampling
      - nonresponse (unit)
      - poststratification
      - they're also trimmed to avoid extremes
      - weight = 0 for 5% of respondents (people not in scope for survery but living house with someone who is)
      - distribution of weights:
        0: 5%
        425: minimum
        9000: average 
        64,000: maximum
  
```{r}
library(survey)
library(foreign)

meps.to.csv <- function(){
  dt <- read.xport('/Users/eczech/data/meps/full_year_consolidated/2013/data.ssp')
  write.csv(dt, file='/Users/eczech/data/meps/full_year_consolidated/2013/data.csv', row.names = F)  
}


d.consol <- read.xport('/Users/eczech/data/meps/full_year_consolidated/2013/data.ssp') %>%
  # Add computed sum of outpatient visits (see http://meps.ahrq.gov/mepsweb/data_stats/download_data/pufs/h163/h163doc.shtml section on outpatient visits for why this is necessary for number of reported outpatient events to equal count in consolidated file)
  mutate(OPDRV13_OPOTHV13=OPDRV13 + OPOTHV13) 

d.util.in <- read.xport('/Users/eczech/data/meps/inpatient_stays/2013/data.ssp')
d.util.er <- read.xport('/Users/eczech/data/meps/er_visits/2013/data.ssp')
d.util.op <- read.xport('/Users/eczech/data/meps/outpatient_stays/2013/data.ssp')
d.util.rx <- read.xport('/Users/eczech/data/meps/prescriptions/2013/data.ssp')


d.long <- read.xport('/Users/eczech/data/meps/longitudinal/2013/data.ssp')
#d.long[,c('OPTOTVY1', 'OPTOTVY2')] %>% 
d.long[,c('RXTOTY1', 'RXTOTY2')] %>% 
  setNames(c('Y1', 'Y2')) %>%
  ggplot(aes(x=Y1, y=Y2)) + geom_jitter()

d.long[,c('ERTOTY1', 'ERTOTY2')] %>% 
  setNames(c('Y1', 'Y2')) %>%
  group_by(Y1, Y2) %>% tally %>%
  ggplot(aes(x=Y1, y=Y2, fill=asinh(n))) + geom_tile()

get.counts <- function(d.consol, d.event, c.event){
  d.consol <- d.consol %>% 
    mutate(DUPERSID=as.character(DUPERSID)) %>% 
    select(DUPERSID, one_of(c.event)) %>%
    rename_('CONSOLIDATED_COUNT'=c.event)
  d.event %>% 
    mutate(DUPERSID=as.character(DUPERSID)) %>% 
    group_by(DUPERSID) %>% tally %>% ungroup %>%
    right_join(d.consol, by='DUPERSID') %>%
    mutate(n = ifelse(is.na(n), 0, n)) %>%
    rename(EVENT_COUNT=n)
}

# Validate that number of individual events matches count in consolidated file
dt <- get.counts(d.consol, d.util.er, 'ERTOT13') 
table(dt$CONSOLIDATED_COUNT == dt$EVENT_COUNT)
table(dt$CONSOLIDATED_COUNT)

dt <- get.counts(d.consol, d.util.in, 'IPDIS13') 
table(dt$CONSOLIDATED_COUNT == dt$EVENT_COUNT)
table(dt$CONSOLIDATED_COUNT)

dt <- get.counts(d.consol, d.util.op, 'OPTOTV13') 
table(dt$CONSOLIDATED_COUNT == dt$EVENT_COUNT)
table(dt$CONSOLIDATED_COUNT)

dt <- get.counts(d.consol, d.util.rx, 'RXTOT13') 
table(dt$CONSOLIDATED_COUNT == dt$EVENT_COUNT)
table(dt$CONSOLIDATED_COUNT)


```


```{r}
d.surv <- d.consol %>% mutate(ONES=1)
d.des <- svydesign(id=~VARPSU, strata=~VARSTR, weights=~PERWT13F, data=d.surv, nest=T)

svytotal(~ONES, d)
svytotal(REGION31~., d)
plot(density(log(d$OBVEXP13 + 1)))

svytotal(~ONES, subset(d.des, AGELAST < 30))
table(d.surv$VARSTR) %>% length
sum(subset(d.surv, AGELAST < 30)$PERWT13F)
```
  