---
title: "CDA"
author: "Scott D Moore"
date: "May 14, 2016"
output: html_document
#output: ioslides_presentation
---
  
# Basic R Training
### Session 1

<hr>
<center><h3>Agenda</h3></center>


- About: Scott_R
- Questions: Why are you here?  What do you think you want to do with R 'Work', 'Play'?
- Notes: rmarkdown will be made available; good, bad, and ugly:).

## Four things about R:

- VERY case sensitive.
- It is fun.
- Somebody has done it or at least thinking about it.
- Many ways to do it RIGHT.

## Today:

1. Work in 'Chunks'.
2. Run 'Swirl'.
3. Get you adjusted to R through self paced.
4. Familiar with some basics.
5. Goal: Be familiar enough so you can follow later presentations. You will not become an expert today.
6. Goal: Be able to navigate so you can start 'producing' today.

**Programming Topics**

- Basic Statistics.
- Data Frames.
- Subsetting.
- Grouping.
- Visualizations (histogram/boxplot/scatterplot/density estimates).
- Goal: Be able to analyze a small problem from data to visual.

<hr>
# Swirl

See the docs here: http://swirlstats.com/students.html

Example:

```
# Install and load swirl
install.packages("swirl")
library(swirl)  

# Install basic programming course
install_from_swirl("R_Programming")

# Run Swirl
swirl()  
```

**"Goodmorning, Wells"**  
  
<hr>

# Programming Examples

R has it own data sets.  

```{r}
data()
head(iris)
```

### Loading New Data

Open a File. Set my Working Directory (Session) to where I want to find my data. 'Pane'  
Read a file. csv is by the far the easist to open strongly recommend this file type for "us".
Remember the "".  

```{r}
cda <- read.csv("~/Downloads/Diabetes.csv", header = TRUE, sep = ",")
head(cda)
```


Alternative table output representations:

<center><strong>Knitr</strong></center>
```{r}
library(knitr)
knitr::kable(head(cda))
```

<center><strong>Datatable (aka DT)</strong></center>
```{r}
library(DT)
DT::datatable(head(cda))
```

<hr>
### Data Summaries and Metadata

Explore the data; object type (character vector), structure, basic statistics.  

You can see what type any symbol is in R with the ```class``` function:
```{r}
class(cda)
```

The ```str``` function will do its best to represent the structure of any object as a string, even complicated multi-dimensional ones with mixed data types:
```{r}
str(cda)
```

Descriptive statistics functions give a quick overview of columns in a data frame (e.g. ```summary```):
```{r}
summary(cda)
```

Other libraries have even more in-depth statistical summaries (e.g. ```psych::describe```):
```{r}
library(psych)
describe(cda)
detach("package:psych", unload=TRUE)
```

<hr>
### Subsetting Data

```{r}
cda.subset <- subset(cda, chol > 125 & weight > 200)
head(cda.subset)
```

Verify that the subset is different from the original:

```{r}
paste('The original dataset had', nrow(cda), 'rows and the subset has only', nrow(cda.subset))
```

"Subsets" can also include "projections", or selections of specific fields:
```{r}
cda.subset <- subset(cda, chol > 125 & weight > 200, select=c('chol', 'weight'))
head(cda.subset)
```


<hr>
### Aggregating Data

"Aggregating" can mean many things, but one simple example would be to group by one variable and create summaries of some kind for another:

```{r}
# Compute mean "glycosolated hemoglobin" for each gender
aggregate(list(glyhb=cda$glyhb),  by=list(gender=cda$gender), FUN=mean) 
```

That didn't work because the variable ```glyhb``` contains NA values (see the summaries above).  To ignore these NA values we can pass ```na.rm=TRUE``` to the aggregate function:

```{r}
aggregate(list(glyhb=cda$glyhb),  by=list(gender=cda$gender), FUN=mean, na.rm=T) 
```

Alternatively, a slightly nicer form of syntax for this is:

```{r}
# Aggregation using R's "formula" operator
aggregate(glyhb ~ gender, cda, FUN=mean, na.action=na.omit)
```

<hr>
### Visualizing Data

<center><em>Univariate Plots</em></center>

Create a Histogram:
```{r}
hist(cda$weight)
```

Similarly, create a density estimate:
```{r}
plot(density(cda$weight, na.rm=T), main='Weight Density Estimate')
```

Create a Boxplot:
```{r}
boxplot(cda$weight, main='Weight Distribution')
```

<center><em>Multivariate Plots</em></center>

Create a Boxplot:
```{r}
boxplot(cda$weight ~ cda$frame, main='Weight by Frame')
```

Create a Scatterplot
```{r}
plot(cda$weight, cda$chol, main='Cholesterol vs Weight')
```


<hr>
# Summary

Your are now able to find your data, create basic statistics, manipulate data, display!

**Homework**: swirl - Basic R Programming.

Thank you. –S
<br><br>
